---
title: "Fly TPC"
author: "Natalie Bright & Chris Terry"
output: 
  html_document: 
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This script compares the fit of different TPC models for the fly experiments. 

```{r message=FALSE, warning=FALSE}
library(rTPC)
library(tidyverse)
library(bbmle)
```

```{r}
Fly_data <- read_csv('Raw Data/Fly counts.csv')
data=Fly_data
```

# Model Building 

## Poisson, No-G0 effect  (1a)

```{r eval=FALSE}
set.seed(1)

Poisson_mod = function(topt, rmax, a){
  ctmax = 30
  Estimate_vec <- rep(NA, nrow(data))
  for(i in 1: nrow(data)){
    temp = data$G1[i]  
    Estimate_vec[i] <- ifelse(temp< topt,
                              rmax * exp(-((temp - topt)/(2 * a))^2), 
                              rmax * (1 - ((temp - topt)/(topt - ctmax))^2))
  }
  
  Estimate_vec[Estimate_vec<0.00001]<-0.00001
  
  return( -sum(dpois( data$Count, Estimate_vec, log = TRUE ))   )
}

Poisson_fit <- mle2(Poisson_mod, 
                    method="L-BFGS-B",
                    upper = c('topt' = 29,'rmax' = 1000,'a' = 100),
                    lower = c('topt' = 20,  'rmax' = 1, 'a' = 0.5),
                    start = list( 'topt' = 25, 'rmax' = 100, 'a' = 1),
                    data = Fly_data)

Poisson_logLik <- logLik(Poisson_fit)
Poisson_logLik
AIC(Poisson_fit)

save(Poisson_fit, file = "Fly TPC models/Poisson_fit")
```

## Negative-Binomial, No-G0 effect (1b)

```{r eval = FALSE}
set.seed(1)

NB_mod = function(topt, rmax, a, shape){
  ctmax = 30
  Estimate_vec <- rep(NA, nrow(data))
  for(i in 1: nrow(data)){
    temp = data$G1[i] 
    
    Estimate_vec[i] <- ifelse(temp< topt,
                              rmax * exp(-((temp - topt)/(2 * a))^2), 
                              rmax * (1 - ((temp - topt)/(topt - ctmax))^2))
  }
  
  Estimate_vec[Estimate_vec<0.00001]<-0.00001
  return( -sum(dnbinom( data$Count,  mu= Estimate_vec,size = 1,log = TRUE ))   )
}

set.seed(1)

NB_AllG0Together_fit <- mle2(NB_mod, 
                             method="L-BFGS-B",
                             upper = c('topt' = 29, 'rmax' = 1000,'a' = 100,
                                       'shape' = 10),
                             lower = c('topt' = 20,'rmax' = 1,  'a' = 0.5,
                                       'shape' = 0.0001),
                             start = list( 'topt' = 25, 'rmax' = 100,
                                           'shape' =1, 'a' = 1),
                             data = Fly_data)

NB_AllG0Together_fit
AIC(NB_AllG0Together_fit)
save(NB_AllG0Together_fit, file = "Fly TPC models/NB_AllG0Together_fit")
```

### Inspection 
```{r}
load("Fly TPC models/NB_AllG0Together_fit" )
Prediction1 <- data.frame( G1 = seq(18, 28, by = 0.01),
                           Expectation = NA)

NB_AllG0Together_fit@coef
ctmax = 30
topt= NB_AllG0Together_fit@coef[1]
rmax= NB_AllG0Together_fit@coef[2]
a= NB_AllG0Together_fit@coef[3]
for(i in 1: nrow(Prediction1)){
  temp = Prediction1$G1[i]  
  Prediction1$Expectation[i] <- ifelse(temp< topt,
                                       rmax * exp(-((temp - topt)/(2 * a))^2), 
                                       rmax * (1 - ((temp - topt)/(topt - ctmax))^2))
}

plot(Prediction1)
```

## Linear dependence of each parameter on temperature (1c)

```{r eval = FALSE}
set.seed(1)

NB_mod_TEMP = function(toptT, topt19, rmaxT, rmax19, aT, a19, shape){
  ctmax = 30
  Estimate_vec <- rep(NA, nrow(data))
  for(i in 1: nrow(data)){
    G0 = data$G0[i] 
    temp = data$G1[i] 
    rmax =   rmax19 + (G0-19) * rmaxT
    topt =  topt19 + (G0-19) * toptT
    a = a19 + (G0-19) * aT 
    Estimate_vec[i] <- ifelse(temp< topt,
                              rmax * exp(-((temp - topt)/(2 * a))^2), 
                              rmax * (1 - ((temp - topt)/(topt - ctmax))^2))
  }
  
  Estimate_vec[Estimate_vec<0.00001]<-0.00001
  
  return( -sum(dnbinom( data$Count,
                        mu= Estimate_vec,
                        size = shape,
                        log = TRUE ))   )
}


data = Fly_data
NB_Linear_Temp_fit <- mle2(NB_mod_TEMP, 
                           method="L-BFGS-B",
                           upper=c('toptT' = 29,'topt19' = 29,
                                   'rmaxT' = 1000,'rmax19' = 1000,
                                   'aT' = 100,'a19' = 100,
                                   'shape' =10),
                           lower=c('toptT' = 20, 'topt19' = 20,
                                   'rmaxT' = 1, 'rmax19' = 1,
                                   'aT' = 0.5, 'a19' = 0.5,
                                   'shape' =0.0001), 
                           start = list( 'toptT' = 0, 'topt19' = 25,
                                         'rmaxT' = 0, 'rmax19' = 100,
                                         'aT' = 0, 'a19' = 1,
                                         'shape' = 1),
                           data = (Fly_data))

NB_Linear_Temp_fit
AIC(NB_Linear_Temp_fit)

save(NB_Linear_Temp_fit, file = "Fly TPC models/NB_Linear_Temp_fit")
```

## Fully split (1d)

```{r eval = FALSE}
set.seed(1)

NB_AllG0Seperate = function(topt19,topt23,topt27, 
                            rmax19,rmax23,rmax27,
                            a19,   a23,   a27, 
                            shape){
  ctmax = 30
  Estimate_vec <- rep(NA, nrow(data))
  
  for(i in 1: nrow(data)){
    G0 = data$G0[i] 
    temp = data$G1[i] 
    rmax =   ifelse(G0==19, rmax19, ifelse(G0==23, rmax23, rmax27))
    topt =   ifelse(G0==19, topt19, ifelse(G0==23, topt23, topt27))
    a =      ifelse(G0==19,    a19, ifelse(G0==23,    a23,    a27))
    Estimate_vec[i] <- ifelse(temp< topt,
                              rmax * exp(-((temp - topt)/(2 * a))^2), 
                              rmax * (1 - ((temp - topt)/(topt - ctmax))^2))
  }
  Estimate_vec[Estimate_vec<0.00001]<-0.00001
  return( -sum(dnbinom( data$Count,
                        mu= Estimate_vec,
                        size = shape,
                        log = TRUE ))   )
}
data = Fly_data

NB_AllG0Seperate_fit <- mle2(NB_AllG0Seperate, 
                             method="L-BFGS-B",
                             upper=c('topt19' = 29,'rmax19' = 1000,
                                     'topt23' = 29,'rmax23' = 1000,
                                     'topt27' = 29, 'rmax27' = 1000,
                                     'shape' =10,
                                     'a19' = 100, 'a23' = 100,
                                     'a27' = 100),
                             lower=c('topt19' = 20,  'rmax19' = 1,
                                     'topt23' = 20, 'rmax23' = 1,
                                     'topt27' = 20, 'rmax27' = 1,
                                     'shape' = 0.0001,
                                     'a19' = 0.5, 'a23' = 0.5,
                                     'a27' = 0.5), 
                             start = list( 'topt19' = 25, 'rmax19' = 100,
                                           'topt23' = 25, 'rmax23' = 100,
                                           'topt27' = 25,'rmax27' = 100,
                                           'shape' =1,
                                           'a19' = 2, 'a23' = 2,
                                           'a27' = 2),
                             data = Fly_data)

NB_AllG0Seperate_fit
AIC(NB_AllG0Seperate_fit)

save(NB_AllG0Seperate_fit, file = "Fly TPC models/NB_AllG0Seperate_fit")
```

## All except `topt` (1e)

```{r eval = FALSE}
set.seed(1)

NB_All_Ex_topt = function(topt, 
                          rmax19,rmax23,rmax27,
                          a19,   a23,   a27, 
                          shape){
  ctmax = 30
  Estimate_vec <- rep(NA, nrow(data))
  
  for(i in 1: nrow(data)){
    G0 = data$G0[i] 
    temp = data$G1[i] 
    rmax =   ifelse(G0==19, rmax19, ifelse(G0==23, rmax23, rmax27))
    #    topt =   ifelse(G0==19, topt19, ifelse(G0==23, topt23, topt27))
    a =      ifelse(G0==19,    a19, ifelse(G0==23,    a23,    a27))
    Estimate_vec[i] <- ifelse(temp< topt,
                              rmax * exp(-((temp - topt)/(2 * a))^2), 
                              rmax * (1 - ((temp - topt)/(topt - ctmax))^2))
  }
  
  Estimate_vec[Estimate_vec<0.00001]<-0.00001
  
  return( -sum(dnbinom( data$Count,
                        mu= Estimate_vec,
                        size = shape,
                        log = TRUE ))   )
}


NB_All_Ex_topt_fit <- mle2(NB_All_Ex_topt, 
                           method="L-BFGS-B",
                           upper=c('rmax19' = 1000, 'rmax23' = 1000,
                                   'rmax27' = 1000,
                                   'shape' =10,
                                   'a19' = 100, 'a23' = 100,
                                   'a27' = 100),
                           lower=c('rmax19' = 1,'rmax23' = 1,
                                   'rmax27' = 1,
                                   'shape' = 0.0001,
                                   'a19' = 0.5, 'a23' = 0.5,
                                   'a27' = 0.5),
                           start = list( 'topt' = 25,
                                         'rmax19' = 100,  'rmax23' = 100,
                                         'rmax27' = 100,
                                         'shape' =1,
                                         'a19' = 1, 'a23' = 1,
                                         'a27' = 1),
                           data = Fly_data)

NB_All_Ex_topt_fit

save(NB_All_Ex_topt_fit, file = "Fly TPC models/NB_All_Ex_topt_fit")
```


## All except `Rmax` (1f)

```{r eval = FALSE}
set.seed(1)

NB_All_Ex_rmax = function(topt19, topt23, topt27,
                          rmax,
                          a19,   a23,   a27, 
                          shape){
  ctmax = 30
  Estimate_vec <- rep(NA, nrow(data))
  
  for(i in 1: nrow(data)){
    G0 = data$G0[i] 
    temp = data$G1[i] 
    # rmax =   ifelse(G0==19, rmax19, ifelse(G0==23, rmax23, rmax27))
    topt =   ifelse(G0==19, topt19, ifelse(G0==23, topt23, topt27))
    a =      ifelse(G0==19,    a19, ifelse(G0==23,    a23,    a27))
    Estimate_vec[i] <- ifelse(temp< topt,
                              rmax * exp(-((temp - topt)/(2 * a))^2), 
                              rmax * (1 - ((temp - topt)/(topt - ctmax))^2))
  }
  
  Estimate_vec[Estimate_vec<0.00001]<-0.00001
  
  return( -sum(dnbinom( data$Count,
                        mu= Estimate_vec,
                        size = shape,
                        log = TRUE ))   )
}
data = Fly_data

NB_All_Ex_rmax_fit <- mle2(NB_All_Ex_rmax, 
                           method="L-BFGS-B",
                           upper=c('topt19' = 29, 'topt23' = 29,  'topt27' = 29,
                                   'shape' =10,
                                   'rmax' = 1000,
                                   'a19' = 100, 'a23' = 100, 'a27' = 100),
                           lower=c('topt19' = 20, 'topt23' = 20, 'topt27' = 20,
                                   'shape' = 0.0001,
                                   'rmax' = 1,
                                   'a19' = 0.5, 'a23' = 0.5, 'a27' = 0.5),
                           start = list( 'topt19' = 25,  'topt23' = 25,  'topt27' = 25,
                                         'shape' =1,
                                         'rmax' = 100,
                                         'a19' = 1, 'a23' = 1,  'a27' = 1),
                           data = Fly_data)

NB_All_Ex_rmax_fit
AIC(NB_All_Ex_rmax_fit)

save(NB_All_Ex_rmax_fit, file = "Fly TPC models/NB_All_Ex_rmax_fit")
```

## All except `a` (1g)

```{r eval = FALSE}
set.seed(1)

NB_All_Ex_a = function(topt19,topt23,topt27, 
                       rmax19, rmax23, rmax27,
                       a, 
                       shape){
  ctmax = 30
  Estimate_vec <- rep(NA, nrow(data))
  
  for(i in 1: nrow(data)){
    G0 = data$G0[i] 
    temp = data$G1[i] 
    rmax =   ifelse(G0==19, rmax19, ifelse(G0==23, rmax23, rmax27))
    topt =   ifelse(G0==19, topt19, ifelse(G0==23, topt23, topt27))
    # a =      ifelse(G0==19,    a19, ifelse(G0==23,    a23,    a27))
    Estimate_vec[i] <- ifelse(temp< topt,
                              rmax * exp(-((temp - topt)/(2 * a))^2), 
                              rmax * (1 - ((temp - topt)/(topt - ctmax))^2))
  }
  
  Estimate_vec[Estimate_vec<0.00001]<-0.00001
  
  return( -sum(dnbinom( data$Count,
                        mu= Estimate_vec,
                        size = shape,
                        log = TRUE ))   )
}
data = Fly_data

NB_All_Ex_a_fit <- mle2(NB_All_Ex_a , 
                        method="L-BFGS-B",
                        upper=c('topt19' = 29,  'rmax19' = 1000,'topt23' = 29,
                                'rmax23' = 1000, 'topt27' = 29,'rmax27' = 1000,
                                'shape' =10,
                                'a' = 100),
                        lower=c('topt19' = 20, 'rmax19' = 1,'topt23' = 20,
                                'rmax23' = 1, 'topt27' = 20,  'rmax27' = 1,
                                'shape' = 0.0001, 
                                a = 0.5),
                        start = list( 'topt19' = 25, 'rmax19' = 100,
                                      'rmax23' = 100, 'rmax27' = 100,
                                      'topt23' = 25, 'topt27' = 25,
                                      'shape' =1,
                                      'a'=1),
                        data = Fly_data)

NB_All_Ex_a_fit
AIC(NB_All_Ex_a_fit)

save(NB_All_Ex_a_fit, file = "Fly TPC models/NB_All_Ex_a_fit")
```


# AIC Results Table

```{r}
load('Fly TPC models/Poisson_fit')
load('Fly TPC models/NB_Linear_Temp_fit')
load('Fly TPC models/NB_AllG0Together_fit')
load('Fly TPC models/NB_AllG0Seperate_fit')
load('Fly TPC models/NB_All_Ex_topt_fit')
load('Fly TPC models/NB_All_Ex_rmax_fit')
load('Fly TPC models/NB_All_Ex_a_fit')
```


```{r}
data_table1 <- data.frame(
  Model = c('1a',
            '1b',
            '1c',
            '1d',
            '1e',
            '1f',
            '1g'),
  Distribution = c('Poisson',
                   'Negative Binomial',
                   'Negative Binomial',
                   'Negative Binomial',
                   'Negative Binomial',
                   'Negative Binomial',
                   'Negative Binomial'),
  Temperature_dependence_of_parameters_Topt_Rmax_a = c('None',
                                                       'None',
                                                       'theta_i=Ttheta_T,i + theta_19,i',
                                                       'theta_i= { If T=19, t ',
                                                       'Like 1d for Rmax and a, and like 1b for Topt',
                                                       'Like 1d for Topt and a, and like 1b for Rmax',
                                                       'Like 1d for Topt and Rmax, and like 1b for a'),
  n = c(3, 4, 7, 10, 8, 8, 8),
  Likelihood = round(c(logLik(Poisson_fit), 
                       logLik(NB_AllG0Together_fit), 
                       logLik(NB_Linear_Temp_fit), 
                       logLik(NB_AllG0Seperate_fit), 
                       logLik(NB_All_Ex_topt_fit), 
                       logLik(NB_All_Ex_rmax_fit), 
                       logLik(NB_All_Ex_a_fit)), digits=2),
  AIC_Difference = round(c((AIC(Poisson_fit)-AIC(NB_AllG0Seperate_fit)), 
                           (AIC(NB_AllG0Together_fit)-AIC(NB_AllG0Seperate_fit)), 
                           (AIC(NB_Linear_Temp_fit)-AIC(NB_AllG0Seperate_fit)), 
                           (AIC(NB_AllG0Seperate_fit)-AIC(NB_AllG0Seperate_fit)), 
                           (AIC(NB_All_Ex_topt_fit)-AIC(NB_AllG0Seperate_fit)), 
                           (AIC(NB_All_Ex_rmax_fit)-AIC(NB_AllG0Seperate_fit)), 
                           (AIC(NB_All_Ex_a_fit)-AIC(NB_AllG0Seperate_fit))), digits=2)
)

data_table1
```

```{r}
write.csv(data_table1, "Outputs/fly model output1.csv", row.names = FALSE)
```

# Parameter summary table

```{r}
parameter_table <- data.frame(
  Parameter = c('Topt', 
                'Rmax',
                'CTmax',
                'a'),
  Description = c('Optimum temperature', 
                  'Maximum reproduction at optimum temperature', 
                  'Critical thermal maximum', 
                  'Related to full curve width'))

parameter_table
write.csv(parameter_table, "parameter_output.csv", row.names = FALSE)

fly_parameter_estimates <- round(NB_AllG0Seperate_fit@coef, digits=2)
fly_parameter_estimates
```

## Confidence Intervals 

Trying to do these separately has just too many dimensions, so assessing CI separately for each of the G0 temperatures

```{r}
NB_mod_CI = function(topt, rmax, a){
  ctmax = 30
  Estimate_vec <- rep(NA, nrow(data))
  for(i in 1: nrow(data)){
    temp = data$G1[i] 
    Estimate_vec[i] <- ifelse(temp< topt,
                              rmax * exp(-((temp - topt)/(2 * a))^2), 
                              rmax * (1 - ((temp - topt)/(topt - ctmax))^2))
  }
  Estimate_vec[Estimate_vec<0.00001]<-0.00001
  
  return( -sum(dnbinom( data$Count, mu= Estimate_vec,
                        size = 1, log = TRUE ))   )
}

set.seed(1)

data = filter(Fly_data, G0 == 19)
NB_G019 <- mle2(NB_mod_CI, 
                data = filter(Fly_data, G0 == 19),
                method="L-BFGS-B",
                upper = c('topt' = 29,'rmax' = 1000,'a' = 100),
                lower = c('topt' = 20,'rmax' = 1, 'a' = 0.5),
                start = list( 'topt' = 25,'rmax' = 100, 'a' = 1) )

NB_G019
```

```{r message=FALSE}
conf_intervals_19 <- confint(NB_G019)

new_row_names_19 <- c("topt19", "rmax19", "a19")
row.names(conf_intervals_19) <- new_row_names_19
```


```{r}
set.seed(1)

data = filter(Fly_data, G0 == 23)
NB_G023 <- mle2(NB_mod_CI, 
                data = filter(Fly_data, G0 == 23),
                method="L-BFGS-B",
                upper = c('topt' = 29,'rmax' = 1000,'a' = 100),
                lower = c('topt' = 20, 'rmax' = 1, 'a' = 0.5),
                start = list( 'topt' = 25, 'rmax' = 100, 'a' = 1) )

NB_G023
```

```{r message=FALSE}
conf_intervals_23 <- confint(NB_G023)

new_row_names_23 <- c("topt23", "rmax23", "a23")
row.names(conf_intervals_23) <- new_row_names_23
```


```{r}
set.seed(1)

data = filter(Fly_data, G0 == 27)
NB_G027 <- mle2(NB_mod_CI, 
                data = filter(Fly_data, G0 == 27),
                method="L-BFGS-B",
                upper = c('topt' = 29, 'rmax' = 1000,  'a' = 100),
                lower = c('topt' = 20,  'rmax' = 1,'a' = 0.5),
                start = list( 'topt' = 25,  'rmax' = 100, 'a' = 1) )

NB_G027
```                      

```{r message=FALSE}
conf_intervals_27 <- confint(NB_G027)

new_row_names_27 <- c("topt27", "rmax27", "a27")
row.names(conf_intervals_27) <- new_row_names_27
```

```{r}

conf_intervals_merged <- rbind(conf_intervals_19, conf_intervals_23, conf_intervals_27)
conf_intervals_merged <- round(conf_intervals_merged, digits = 2)
conf_intervals_merged
```

# Predictions

```{r}
new_data_seq <- data.frame(G1 = seq(18, 28, by = 0.1))
Prediction <- data.frame(G1 = rep(new_data_seq$G1, 3),
                         G0 = rep(c(19, 23, 27), each = nrow(new_data_seq)), Expectation=NA)

NB_AllG0Seperate_fit@coef
ctmax = 30
topt19= NB_AllG0Seperate_fit@coef[1]
topt23= NB_AllG0Seperate_fit@coef[2]
topt27= NB_AllG0Seperate_fit@coef[3]
rmax19= NB_AllG0Seperate_fit@coef[4]
rmax23= NB_AllG0Seperate_fit@coef[5]
rmax27= NB_AllG0Seperate_fit@coef[6]
a19= NB_AllG0Seperate_fit@coef[7]
a23= NB_AllG0Seperate_fit@coef[8]
a27= NB_AllG0Seperate_fit@coef[9]

for(i in 1: nrow(Prediction)){
  rmax =   ifelse(Prediction$G0[i]==19, rmax19, 
                  ifelse(Prediction$G0[i]==23, rmax23, rmax27))
  topt =   ifelse(Prediction$G0[i]==19, topt19, 
                  ifelse(Prediction$G0[i]==23, topt23, topt27))
  a =      ifelse(Prediction$G0[i]==19,    a19, 
                  ifelse(Prediction$G0[i]==23,    a23,    a27)) 

  temp = Prediction$G1[i]  
  Prediction$Expectation[i] <- ifelse(temp< topt,
                                      rmax * exp(-((temp - topt)/(2 * a))^2), 
                                      rmax * (1 - ((temp - topt)/(topt - ctmax))^2)) 
  }
```

## Confidence interval around model. 


```{r}
SHAPE <- NB_AllG0Seperate_fit@coef['shape']

Prediction$LowerExp <-qnbinom(p = 1/6,size = SHAPE, mu = Prediction$Expectation )
Prediction$UpperExp <-qnbinom(p = 5/6,size = SHAPE, mu = Prediction$Expectation )

qnorm( 0.975, mean =  0, sd = 1)



Prediction


```



## Plotting

```{r}

Fly_data %>%
 rename ( `G0 Temperature` = G0) %>%
  ggplot() +
  geom_jitter(aes(G1, Count), alpha=0.5, size = 1, width = 0.15, height =0) +
  geom_line(aes(G1, Expectation), col = '#0072B2',
            data =  rename (Prediction,`G0 Temperature` = G0))+
  theme_bw(base_size = 12) +
  theme(text = element_text(size = 10),
        strip.background = element_rect(fill =NA)) +
  scale_x_continuous(breaks = seq(19, 27, by = 2)) +
  labs(x = 'G1 Temperature (ºC)',
       y = 'Emerged Offspring')+
  facet_wrap(~`G0 Temperature`, labeller = label_both )

  ggsave('Figures/Fly TPC Fit.png', width = 6, height = 4 )
  
Fly_data %>%
  group_by(  G0,G1)%>% 
  summarise(MEAN= mean(Count),
            SE = sd(Count)/sqrt(n())  ) %>%
  rename ( `G0 Temperature` = G0) %>%
  ggplot() +
  geom_point(aes(G1, MEAN), size = 1) +
    geom_point(aes(G1, Count), alpha=0.5, size = 0.5, data = rename (Fly_data, `G0 Temperature` = G0) ) +
  geom_errorbar(aes( x = G1, ymin = MEAN-SE, ymax = MEAN+SE))+
  geom_line(aes(G1, Expectation), col = '#0072B2',
            data =  rename (Prediction,`G0 Temperature` = G0) ) +
  theme_bw(base_size = 12) +
  theme(text = element_text(size = 10),
        strip.background = element_rect(fill =NA)) +
  scale_x_continuous(breaks = seq(19, 27, by = 2)) +
  labs(x = 'G1 Temperature (ºC)',
       y = 'Emerged Offspring')+
  facet_wrap(~`G0 Temperature`, labeller = label_both )

ggsave('Figures/Fly TPC Fit_se.png', width = 6, height = 4 )

```


# Session Info
```{r}
sessionInfo()
```

