---
title: "Simulations"
author: "Natalie Bright & Chris Terry"
output: 
  html_document: 
    toc: true
editor_options: 
  chunk_output_type: console
---

# Introduction 

This script conducts and analyses the simulations. The simulation scripts themselves are stored separately `Scripts/`. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(bbmle)
library(viridis)
library(RColorBrewer)
library(wesanderson)
library(psych) ## for geometric mean
library(lme4)
library(lmerTest)
library(cowplot)
library(knitr)
```


# Fly interpolation function


## Parameters and interpolation

```{r}
load('Fly TPC models/NB_AllG0Seperate_fit')

topt19  = 28.3656893
topt23 =  28.3467465
topt27 =  29.0000000
rmax19 =  46.1107868 
rmax23 =  71.3059568 
rmax27 =  20.4986572 
a19   = 4.2125259  
a23   = 3.6526915 
a27   = 3.2156789 

cTmax = 30

## Rmax

Rmax_interp<- Vectorize(function(G0){
  Rmax<- ifelse( G0<23, rmax19 + ((G0-19)/4) * (rmax23 - rmax19),
                 rmax23 + ((G0-23)/4) * (rmax27 - rmax23))
  ifelse( Rmax< 0.1, return(0.1), return(Rmax))
})

## Topt
Topt_interp<- Vectorize(function(G0){
  Topt<-  ifelse( G0<23, topt19 + ((G0-19)/4) * (topt23 - topt19),
                  topt23 + ((G0-23)/4) * (topt27 - topt23)) 
  
  ifelse( Topt>(cTmax-0.1) , return(cTmax-0.1), return(Topt))
})

## A
a_interp<- Vectorize(function(G0){
  ifelse( G0<23, a19 + ((G0-19)/4) * (a23 - a19),
          a23 + ((G0-23)/4) * (a27 - a23))
})
```

## Plotting interpolations

```{r}
## checking it works
G0_seq <- seq( from = 19, to = 30, by =0.1)

data.frame( G0= seq( from = 19, to = 30, by =0.1))%>%
  mutate( Topt = Topt_interp(G0), 
          Rmax= Rmax_interp(G0),
          a = a_interp(G0) ,
          Col = ifelse( G0 %in% c(19,23,27), 'red', 'black')) %>%
  pivot_longer(cols = c(Topt,Rmax, a ),
               names_to = 'Parameter',  values_to = 'Value') %>%
  ggplot(aes( x = G0, y = Value, col = Col))+
  geom_point()+
  facet_wrap(~Parameter, scale = 'free_y', labeller = label_both)+
  scale_color_identity()+
  scale_x_continuous(breaks = 19:30)+
  xlab('G0 Temperature (°C)')+
  ylab('')+
  ggtitle('Fly Parameter Interpolation')+
  theme_bw() -> EPCFigA

EPCFigA
```

## Function to calculate Environmental Performance Curve

```{r}
EPC_trans = function(G1_temp, G0_temp, cTmax){
  Rmax<- Rmax_interp(G0_temp)
  Topt<- Topt_interp(G0_temp)
  a   <- a_interp(   G0_temp)
  R =  Rmax* ifelse(G1_temp < Topt,
                    exp(-((G1_temp - Topt)/(2 * a))^2), 
                    1 - ((G1_temp - Topt)/(Topt - cTmax))^2)
  return(R)
}
```

## Plotting full EPCs

```{r}
color_palette1 <- wes_palette("Zissou1", 12, type = "continuous")

expand_grid(G0 = seq(from = 19, to = 30, by = 1),
            G1 = seq(from = 19, to = 30, by = 0.01 )) %>%
  mutate( Rmax = Rmax_interp(G0),
          Topt = Topt_interp(G0),
          a = a_interp(G0), 
          R_predict = EPC_trans( G1_temp = G1, G0, cTmax=30)) %>%
  mutate(R_predict = ifelse(R_predict<0, 0, R_predict )) %>%
  ggplot(aes(x = G1, y = R_predict, col = factor(G0))) +
  geom_line() +
  labs(title="Fly Environmental Performance Curve",
       x="G1 Temperature (°C)", y="Population growth rate",
       colour="G0 Temperature\n(°C)") +
  guides(col=guide_legend(ncol=2))+
  scale_color_manual(values = color_palette1) +
  theme_bw()-> EPCFigB

EPCFigB

XXXX<- plot_grid(EPCFigA, EPCFigB, ncol = 1,
          rel_heights = c(1, 1),
          labels = c('a)', 'b)'))

ggsave(plot = XXXX, filename ='Figures/FlyParamPlots.png', height =6, width =7)


```

# Utility Functions and parameters

```{r}
setup<- list( 'Av_temp' = 24, 
              'MaxTemp' = 29, 
              'burnin' = 500,  #burn-in time
              'CC_time' = 2000, #climate change time - 2000 gens = so to around year 2100 - at worst climate change scenarios could be up to 4.7 degrees of warming above baseline by 2100 (IPCC report)
              'temp_sd' = 1,   #standard deviation of temp - sd of Australian data summer av mid temps = 1.1
              'Start_flies' = 10000, 
              'Start_wasps' = 100, 
              'threshold' = 1, #below 1 individual, deemed extinct
              'AutoCorrelated' = FALSE,
              'autocorrelation' = 0.8, #ac of Australian av mid temp = 0.879
              'transgenerational_fly'= TRUE,
              'transgenerational_wasp' = TRUE) 

params<- list(
  ## fly params
  'topt'= 26.313093,  ## from all together fit 
  'rmax' =  38.854392, ## from all together fit 
  'fly_a' = 3.052021, ## from all together fit
  'ctmax' = 30, 
  'fly_K' = 100000,     #setting large carrying capacity 
  'fly_fail'= 20,
  ## wasp params
  'wasp_aMODA' = 0.14*0.00026,
  'wasp_aMODB' = 0.00026,
  'wasp_h' = 0.01, ## set like this because 1/100 = max number of eggs
  'wasp_k' = 0.65 
)

```

# Simulations

```{r}
## Load models fitted to experimental data
load('Outputs/WaspModel') ##waspglm2_poly
load('Outputs/WaspModel_NoG0') ##waspglm2_no_G0
##NB model predictions will be on generalised scale use boot::inv.logit() to back transform 

## Load simulation model functions
source('Scripts/TestExtinctionPoint_ModelA.R' )
source('Scripts/TestExtinctionPoint_ModelB.R' )

```

## Generating Temperature Sequences

```{r}
L <- setup$burnin + setup$CC_time  #overall time
set.seed(1)
Env_vec <- c( rep(setup$Av_temp,setup$burnin), 
              seq( from = setup$Av_temp, to = setup$MaxTemp, 
                   length.out = setup$CC_time )) + rnorm(L,  mean = 0,
                                                         sd = setup$temp_sd)

Env_vec[ Env_vec>(params$ctmax -0.1) ] <- (params$ctmax-0.1) #stop going above ctmax
```

## Autocorrelated sequences

```{R fig.height=4, fig.width=4}
source('Scripts/Make_AC_data.R')
## testing:
Make_AC_data(2000, sd=3.567, AC=0.8) -> X
plot(X[1:1999]~X[2:2000])
mean(X)
sd(X)
```

## Testing Examples

```{r fig.height=3}
setup$transgenerational_fly<-TRUE
setup$transgenerational_wasp<-TRUE

TestExtinctionPoint_ModelB(3, params = params, setup=setup, PLOT=TRUE )
ggsave('Figures/TrajectoryExamples.png', width =10, height = 4, bg = 'white', dpi = 1000)

setup$transgenerational_fly<-FALSE
setup$transgenerational_wasp<-FALSE
TestExtinctionPoint_ModelB(1, params = params, setup=setup, PLOT=TRUE )
TestExtinctionPoint_ModelA(1, params = params, setup=setup, PLOT=TRUE )

```

# Main tests

```{R}
NUM_REPS = 100
SimResults <-data.frame( TrialRep = 1:NUM_REPS )

```

## Model A 

```{r eval = FALSE}
## No autocorrelation

setup$AutoCorrelated<-FALSE
setup$transgenerational_fly<-TRUE
setup$transgenerational_wasp<-TRUE

SimResults$ModelA_NoAuto_BothTrans <- map_int(.x = 1:NUM_REPS,
                                              .f =  TestExtinctionPoint_ModelA,
                                              params = params, setup=setup, PLOT=FALSE)

setup$transgenerational_fly<-FALSE
setup$transgenerational_wasp<-FALSE
SimResults$ModelA_NoAuto_NoTrans <- map_int(.x = 1:NUM_REPS,
                                            .f =  TestExtinctionPoint_ModelA,
                                            params = params, setup=setup, PLOT=FALSE)

## With autocorrelation
setup$AutoCorrelated<-TRUE
setup$transgenerational_fly<-TRUE
setup$transgenerational_wasp<-TRUE
SimResults$ModelA_Auto_BothTrans <- map_int(.x = 1:NUM_REPS,
                                            .f =  TestExtinctionPoint_ModelA,
                                            params = params, setup=setup, PLOT=FALSE)

setup$transgenerational_fly<-FALSE
setup$transgenerational_wasp<-FALSE
SimResults$ModelA_Auto_NoTrans <- map_int(.x = 1:NUM_REPS,
                                          .f =  TestExtinctionPoint_ModelA,
                                          params = params, setup=setup, PLOT=FALSE)

```

## Model B

```{r eval = FALSE}
## No autocorrelation
setup$AutoCorrelated<-FALSE
setup$transgenerational_fly<-TRUE
setup$transgenerational_wasp<-TRUE

SimResults$ModelB_NoAuto_BothTrans <- map_int(.x = 1:NUM_REPS,
                                              .f =  TestExtinctionPoint_ModelB,
                                              params = params, setup=setup, PLOT=FALSE)

setup$transgenerational_fly<-FALSE
setup$transgenerational_wasp<-FALSE
SimResults$ModelB_NoAuto_NoTrans <- map_int(.x = 1:NUM_REPS,
                                            .f =  TestExtinctionPoint_ModelB,
                                            params = params, setup=setup, PLOT=FALSE)

## With autocorrelation
setup$AutoCorrelated<-TRUE
setup$transgenerational_fly<-TRUE
setup$transgenerational_wasp<-TRUE
SimResults$ModelB_Auto_BothTrans <- map_int(.x = 1:NUM_REPS,
                                            .f =  TestExtinctionPoint_ModelB,
                                            params = params, setup=setup, PLOT=FALSE)

setup$transgenerational_fly<-FALSE
setup$transgenerational_wasp<-FALSE
SimResults$ModelB_Auto_NoTrans <- map_int(.x = 1:NUM_REPS,
                                          .f =  TestExtinctionPoint_ModelB,
                                          params = params, setup=setup, PLOT=FALSE)

```

## Adding one on, one off transgenerational effects


```{r eval = FALSE}
### Model A 
## No autocorrelation
setup$AutoCorrelated<-FALSE
setup$transgenerational_fly<-FALSE
setup$transgenerational_wasp<-TRUE

SimResults$ModelA_NoAuto_WaspTrans <- map_int(.x = 1:NUM_REPS,
                                              .f =  TestExtinctionPoint_ModelA,
                                              params = params, setup=setup, PLOT=FALSE)

setup$transgenerational_fly<-TRUE
setup$transgenerational_wasp<-FALSE
SimResults$ModelA_NoAuto_FlyTrans <- map_int(.x = 1:NUM_REPS,
                                            .f =  TestExtinctionPoint_ModelA,
                                            params = params, setup=setup, PLOT=FALSE)

## With autocorrelation
setup$AutoCorrelated<-TRUE
setup$transgenerational_fly<-FALSE
setup$transgenerational_wasp<-TRUE
SimResults$ModelA_Auto_WaspTrans <- map_int(.x = 1:NUM_REPS,
                                            .f =  TestExtinctionPoint_ModelA,
                                            params = params, setup=setup, PLOT=FALSE)

setup$transgenerational_fly<-TRUE
setup$transgenerational_wasp<-FALSE
SimResults$ModelA_Auto_FlyTrans <- map_int(.x = 1:NUM_REPS,
                                          .f =  TestExtinctionPoint_ModelA,
                                          params = params, setup=setup, PLOT=FALSE)
### Model B
## No autocorrelation
setup$AutoCorrelated<-FALSE
setup$transgenerational_fly<-FALSE
setup$transgenerational_wasp<-TRUE

SimResults$ModelB_NoAuto_WaspTrans <- map_int(.x = 1:NUM_REPS,
                                              .f =  TestExtinctionPoint_ModelB,
                                              params = params, setup=setup, PLOT=FALSE)

setup$transgenerational_fly<-TRUE
setup$transgenerational_wasp<-FALSE
SimResults$ModelB_NoAuto_FlyTrans <- map_int(.x = 1:NUM_REPS,
                                            .f =  TestExtinctionPoint_ModelB,
                                            params = params, setup=setup, PLOT=FALSE)

## With autocorrelation
setup$AutoCorrelated<-TRUE
setup$transgenerational_fly<-FALSE
setup$transgenerational_wasp<-TRUE
SimResults$ModelB_Auto_WaspTrans <- map_int(.x = 1:NUM_REPS,
                                            .f =  TestExtinctionPoint_ModelB,
                                            params = params, setup=setup, PLOT=FALSE)

setup$transgenerational_fly<-TRUE
setup$transgenerational_wasp<-FALSE
SimResults$ModelB_Auto_FlyTrans <- map_int(.x = 1:NUM_REPS,
                                          .f =  TestExtinctionPoint_ModelB,
                                          params = params, setup=setup, PLOT=FALSE)

```

## Combining and saving

```{r eval = FALSE}
SimResults %>%
  pivot_longer(cols= -TrialRep,
               names_to = 'Name', values_to = 'Extinctions') %>%
  separate(Name, into = c('Model', 'AC', 'TransGen')) -> LongFormExtData

LongFormExtData

write_csv(LongFormExtData, "Outputs/CombinedLongSimulationOutputs.csv")

```


# Simulation Results Analysis

```{r}
LongFormExtData <- read_csv("Outputs/CombinedLongSimulationOutputs.csv")
```

```{r}
LongFormExtData %>%
  group_by(Model, AC, TransGen) %>%
  summarise(GeomMean = psych::geometric.mean(Extinctions)) %>%
  pivot_wider(id_cols = c(AC, TransGen),
              names_from = Model,
              values_from = GeomMean) %>% 
  knitr::kable(digits = 2)



LongFormExtData %>%
  #filter(TransGen %in% c("BothTrans","NoTrans" ) ) %>%
  ggplot( aes( x = Extinctions))+
  geom_density(fill="#0072B2", aes( fill ="black"))   +
#  geom_histogram(binwidth=log10(50), fill="#0072B2", colour="black")+
  facet_grid(Model+AC~TransGen,
             labeller = labeller(TransGen=c("BothTrans" = "Transgenerational Effects", 
                                            "NoTrans" = "No Transgenerational Effects"),
                                 AC=c("Auto" = "Autocorrelation",
                                      "NoAuto" = "No Autocorrelation"),
                                 Model=c("ModelA" = "Model A",
                                         "ModelB" = "Model B")), switch = 'y')+
   theme_bw(base_size = 12) +
  theme(text = element_text(size = 10),
        strip.background = element_rect(fill =NA)) +
  labs(x = "Time to extinction (generations)", y = "Frequency")+
  scale_x_log10(limits = c(200, 3000))



LongFormExtData %>%
  mutate(TransMod = case_match(TransGen,
                                           "BothTrans" ~ "Both", 
                                           "NoTrans" ~ "None",
                                           'WaspTrans' ~ "Wasp Only"  ,
                                           'FlyTrans'  ~ "Fly Only")) %>%
  #filter(TransGen %in% c("BothTrans","NoTrans" ) ) %>%
  ggplot( aes( x = Extinctions))+
  geom_density(aes( fill = fct_relevel(TransMod, rev(c("Both", "Wasp Only","Fly Only", "None")))),
                    alpha = 0.5)   +
  #  geom_histogram(binwidth=log10(50), fill="#0072B2", colour="black")+
  facet_grid(AC~Model,
             labeller = labeller(AC=c("Auto" = "With Autocorrelation",
                                      "NoAuto" = "No Autocorrelation"),
                                 Model=c("ModelA" = "Model A",
                                         "ModelB" = "Model B")), switch = 'y')+
  theme_bw(base_size = 12) +
  scale_fill_manual( name = 'Transgenerational\nEffects', values = c('blue','gold','darkgreen', 'red'))+
  theme(text = element_text(size = 10),
        strip.background = element_rect(fill =NA), legend.position = 'bottom') +
  labs(x = "Time to extinction (generations)", y = "Frequency Density")+
  scale_x_log10(limits = c(200, 3000))


ggsave('Figures/SimulationHistograms.png', width = 6, height =5)

```

## Checking for overall interaction

```{r}
## reordering factors 

LongFormExtData %>%
   # filter(TransGen %in% c("BothTrans","NoTrans" ) ) %>%
  mutate( AC = fct_relevel(AC, 'NoAuto'), 
          TransGen = fct_relevel(TransGen, 'NoTrans') ) -> LongFormExtData

## Interaction between trans-generational effects and autocorrelation
model_geomA <- lmer( log10(Extinctions) ~ AC * TransGen +(1|TrialRep),
                     data=filter(LongFormExtData, Model =='ModelA')) 
model_geomB <- lmer( log10(Extinctions) ~ AC * TransGen +(1|TrialRep),
                     data=filter(LongFormExtData, Model =='ModelB')) 


summary(model_geomA) 
summary(model_geomB)

## NB Satterthwaite's method:
anova(model_geomA) 
anova(model_geomB) ## Not significant
```

```{R}
LongFormExtData %>%
  mutate( FlyTrans = TransGen %in% c("BothTrans", 'FlyTrans' ),
          WaspTrans = TransGen %in% c("BothTrans", 'WaspTrans' ) ) -> SplitTrans
```

## Full model (three-way interactions)
```{r}

model_geomA_Full <- lmer( log10(Extinctions) ~ AC*FlyTrans*WaspTrans+(1|TrialRep),
                     data=filter(SplitTrans, Model =='ModelA')) 
model_geomB_Full <- lmer( log10(Extinctions) ~   AC*FlyTrans*WaspTrans+(1|TrialRep),
                     data=filter(SplitTrans, Model =='ModelB')) 

summary(model_geomA_Full)
summary(model_geomA_Full)$coefficients[,1:3] %>% knitr::kable()


summary(model_geomB_Full)
summary(model_geomB_Full)$coefficients[,1:3] %>% knitr::kable()

anova(model_geomA_Full) %>% knitr::kable()
anova(model_geomB_Full) %>% knitr::kable()## Not significant

```

## Only 2-way interactions
```{r}
model_geomA_Split <- lmer( log10(Extinctions) ~ AC + FlyTrans+WaspTrans +AC:FlyTrans  +AC:WaspTrans  + FlyTrans:WaspTrans+(1|TrialRep),
                     data=filter(SplitTrans, Model =='ModelA')) 

model_geomB_Split <- lmer( log10(Extinctions) ~  AC + FlyTrans+WaspTrans +AC:FlyTrans  +AC:WaspTrans  + FlyTrans:WaspTrans+(1|TrialRep),
                     data=filter(SplitTrans, Model =='ModelB')) 

summary(model_geomA_Split) 
summary(model_geomB_Split)
## NB Satterthwaite's method:
anova(model_geomA_Split) 
anova(model_geomB_Split) ## Not significant
```

```{r}
## Model A, isolating effect of fly trans by just having wasptrans
model_geomA_focus <- lmer( log10(Extinctions) ~ AC*FlyTrans+(1|TrialRep),
                     data=filter(SplitTrans, Model =='ModelA', WaspTrans)) 
summary(model_geomA_focus)
```



# Session Info
```{r}
sessionInfo()
```

